Optimal Fitting of Planar Curves to Prescribed Constraints

### 2013-07-19 idp/documentation/paper.tex
continued working on usability section

			\begin{enumerate}
				\item obtain the source curve as either a real or a mental image of the curve
				\item extract a set of properties from the source curve that uniquely identify it
				\item provide these properties to the software
				\item have the software turn them into the result curve
			\end{enumerate}

@@

			Step 1 is of course independent of any software that may be used by the designer. Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it accepts as descriptions for curves (the `language' the software uses). Finally, the success in step 4 depends solely on the ability of the software to turn descriptions of curves that are given to it into mathematical curves.

			The influence which the design tool has on the success or failure of this whole process thus depends on the following criteria.

			\begin{enumerate}
				\item Can the language describe curves succinctly and uniquely?
				\item Is it easy for humans to `read' descriptions in the language from curves?
				\item Is it easy for humans to `speak' descriptions in the language to the software?
				\item Is it possible for software to build curves from descriptions in the language?
			\end{enumerate}

			At this point it seems obvious that the choice of language is the single most influential one when it comes to the usability of curve design software. Ideally, the language should be very expressive, facilitating points 1 to 3, but it should still be possible to fulfill point 4 and derive mathematical curves from a description, a process which becomes increasingly difficult with more expressive description languages. Thus, usability analysis in the subsequent sections will focus almost exclusively on this concept of language and the key points made in the list above.

@@

				% TODO - illustrations
				% curves that are best described in terms of curvature are difficult to put down
				%   an excessive number of control points are needed to pin down curvature
				%     compromises smoothness
				%     makes the curve hard to change
				%     somewhat okay when tracing an existing image, since the shapes are already final and the source is likely smooth
				%   examples
				%     circle arcs (may have too much curvature in the middle, no constant curvature)
				%     spirals (fluttershy's mane)
				%     lugia's back line
				% changind direction at some node usually adversely affects neighboring spline segments
				%   changing the direction on the left side of a node with a long segment to the right screws up the right side
				%     bézier splines do not have much degrees of freedom left after direction continuity is ensured, even though curvature continuity may still be violated
				%     requires additional segments, tradeoff between smoothness and locality, not much tool support for retaining as much smoothness as possible while adding more freedom
				% curvature continuity is not guaranteed
				%   curves with lots of changes in curvature sign are affected the most
				%   examples
				%     lugia's belly

>>>>

			\begin{enumerate}
				\item obtain the source curve as either a real or a mental image of the curve
				\item extract some properties from the source curve
				\item provide these properties to the software
				\item have the software derive the most likely result curve from these properties
			\end{enumerate}

@@

			Step 1 is of course independent of any software that may be used by the designer. Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it accepts as descriptions for curves (the `language' the software uses). Unless the properties the designer conveyed to the software in step 3 uniquely identify the curve (an unlikely situation in curve design), in step 4, the design tool will have to choose the curve that the user was most likely describing from an infinite set of curves that exhibit the supplied properties. This notion of 'most likely' is usually modelled in terms of a fairness measure, fairer curves being regarded as more likely. Fairness ususally encodes intuitive notions such as smoothness using mathematical concepts such as continuity and curvature. Here, the software has an opportunity to influence the design process both positively and negatively depending on how the fairness measure is chosen as well as by how good the software is at reliably finding a curve that has a high fairness.

			At this point it seems like the two ways in which curve design software can affect the design process are through the choice of language and fairness measure. Ideally, the language should be very expressive, enabling the user to communicate their intent to the software efficiently, while it and the fairness measure should still retain the mathematical properties that allow curves to be derived from them, a process which becomes increasingly difficult with more expressive description languages and more complex fairness measures.

			From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools.

			\begin{enumerate}
				\item How well can the language describe curves?
				\item How well can humans to `read' descriptions in the language from curves?
				\item How well can humans to `speak' descriptions in the language to the software?
				\item How well does the fairness measure capture the intuitive notion of smoothness?
				\item How well can the software derive curves from descriptions and fairness?
			\end{enumerate}

@@

				% TODO - illustrations
				% curves that are best described in terms of curvature are difficult to put down
				%   an excessive number of control points are needed to pin down curvature
				%     compromises smoothness
				%     makes the curve hard to change
				%     somewhat okay when tracing an existing image, since the shapes are already final and the source is likely smooth
				%   examples
				%     circle arcs (may have too much curvature in the middle, no constant curvature)
				%     spirals (fluttershy's mane)
				%     lugia's back line
				% curvature continuity is not guaranteed
				%   curves with lots of changes in curvature sign are affected the most
				%   examples
				%     lugia's belly
